<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Filtering in rmangal}
-->

```{r load,warning=FALSE, eval=TRUE, echo=FALSE, message=FALSE}
library(rmangal)
```
# Filtering

The `rmangal` package allows to *filter* ressources according to a series
of pre-defined relationships. Filtering is done using the `mangalSearch`
function. The filters are lists, that have three attributes: `field`,
`relation`, and `target`. The basis of filtering is: the `field` of a given
type of resource will be matched for `target`, by a `releation`.

The `relation` can be any of `contains` (target matches anywhere in the field),
`startswith` or `endswith` (matches at the beginning or end of field), `exact`
(the field value is *exactly* target), `in` (target is one of the values of
the field), `range` (target is between the two bounds), and `gte`, `lte`,
`lt`, and `gt` (for greater/less or equal than).

By defaults, all fields are searchable (with the exception of `id` and
`owner`).

# Examples

```{r connec}
api <- mangalapi()
```

## Looking for all species of a given genus

To look for all species of  the genus `Heteractis`, we write the following
filter:

```{r spFilter}
heteractis <- list(field='name', relation='startswith', target='Heteractis')
Heteractis_species <- mangalSearch(api, 'taxa', list(heteractis))
laply(Heteractis_species, function(x) x$name)
```

Note that the last arguments of `mangalSearch` (`filters`) is of type
`list`. This allows to have multiple filters, as we'll see in the next example.

## Looking for networks within a geographical boundary

We want to find networks within a square defined by its latitude/longitude coordinates. We will define two boundaries: one for the latitudes, and one for the longitudes. This can be done using `range`, and two filters:

```{r filtLatLon}
f_lat <- list(field='latitude', relation='range', target=c(1.756, 1.813))
f_lon <- list(field='longitude', relation='range', target=c(124.760, 124.808))
Networks <- mangalSearch(api, 'network', list(f_lat, f_lon))
laply(Networks, function(x) x$name)
```
