<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Basics of rmangal}
-->

```{r load,warning=FALSE, eval=TRUE, echo=FALSE, message=FALSE}
library(rmangal)
```
# rmangal - R access to hosted MANGAL API

The `mangal` project is a [data specification][dataspec] and [API], desgined
to facilitate the retrieval, archival, and re-use of data on ecological
interactions.

[API]: http://mangal.uqar.ca/api/v1/?format=json
[dataspec]: http://mangal.uqar.ca/doc/spec/

# An overview of the mangal format

The database (and underlying data format) is built around the idea that
networks require meta-data to be fully understood. Instead of presenting
them as 0/1 matrices, the `mangal` data format uses relations between objects
as small as a possible. So as to understand the different elements, let's
connect to a server implementing the API:

```{r firstStep}
netdb <- mangalapi()
```

The `netdb` object has all the necessary informations for the `rmangal`
package to work (and is the first argument of many functions). Let's look
at the names of this object:

```{r methodsOK}
netdb$resources
```

Note that if you are logged-in at this point (see the vignette about
*Contributing data*), `netdb` will gain the attributes `auth` and `me`, so
let's disregard these. Each element in this array corresponds to a *resource*,
*i.e.* a type of object you can interact with. The data specification is
a description of (i) the content of each field, and (ii) how it should be
formatted. As no one is supposed to remember the whole data specification,
a function called `whatIs` will give you a brief overview of what each field
is supposed to mean:

```{r whatIs}
whatIs(netdb, 'taxa')
```

The `rmangal` package will return you objects as `list`s, and (in case you
want to contribute data), will expect objects in the same format. In the
`data.frame` returned by `whatIs`, there are all the informations to understand
the objects that are returned. If you are not interested in contributing data,
you will most likely be OK with the first two columns: the name of the field,
and what it means. For example, this data frame gives you all the meta-data
associated with a taxa. The `null` column will also tell you which fields
are facultative, and which are mandatory.

Let's compare this output with a `taxa` object pulled from the database
(we'll see each element of the data structure and how to access it just after):

```{r showtaxa}
getTaxa(netdb, 9)
```

## Describing nodes

Nodes in the networks can be of type `taxa`, `population`, and `item`. Calling
`whatIs` on `item` or `population` will show that `population`s are linked to a
`taxa`, and that `item`s are linked to a population. Here are the reasons why.

A `population` is, simply put, one occurence of a `taxa`. If you sample
several networks over space, and find the same species at different places,
each of these local occurences is best represented as a `population`, linked
to its `taxa`. If you sample individuals within these populations, or if
you sample the network repeatedly through time, then each individual/time
point is an `item` attached to this population. The `item` objects have a
`level` property, which is either `individual` (invidual-based network) or
`population` (time-replicated sampling). The important point is that, when
looking at an `item` or a `population`, it is possible to move up to the
`taxa` to which they belong.

## Describing interactions

Interactions and networks are represented by `interaction`, `network`, and
`dataset` - quite simply, a `dataset` is an ensemble of `network`s (*i.e.*
at least one), and each `network` is an ensemble of `interaction`s (*i.e.*
at least one). And each `interaction` is a link between (*a minima*) two
`taxa`. The `interaction` format has `*_to` and `*_from` fields for all
level of the data specification representing nodes (`taxa`, `population`,
`item`). It is required that all of the "upper" levels are filled. For example,
if the interaction is at the *individual* level, the `interaction` object will
still have the `taxa_to`, `taxa_from`, `population_to` and `population_from`
fields filled.

## Meta-data

# Pulling data from the database

The data can be accessed with either `get*` or `list*`. A function starting with `get` will retrieve a *single* record, identified by its `id`. A function starting by `list` will return *all* records of a given type (options to filter will be added in future releases). All the functions follow the same naming convention: either `get` or `list`, and the name of the resource (`taxa`, `reference`) with its first letter capitalized. So getting a list of all networks with their name, unique identifier, and number of interactions, is as simple as

```{r listnetwork}
head(ldply(listNetwork(netdb), summarize, id = id, name = name, n_int = length(interactions)))
```

If we want to have a look at  the first network, we simply need to write

```{r getnetwork}
(net1 <- getNetwork(netdb, 2))
```

And similarly, the content of the first interaction of this network is 

```{r firstInterac}
(int1 <- getInteraction(netdb, net1$interactions[1]))
```

To get a sense of what each property mean, you simply need to call `whatIs(netdb, 'interaction')`.

# Example: plotting a network

With this information in hand, getting a full network with all taxa information is simply a matter of following each interaction down to the taxa level, and putting this together in a single object. It's easy, but tedious. The `toIgraph` function will take care of it automatically, and pull a network a an `igraph` object:

```{r, warning=FALSE, fig.cap="A visualisation of the network."}
#G <- toIgraph(netdb, 2)
#plot(G)
```
