<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Uploading data}
-->

```{r load}
library(rmangal)
```

# Contributing data through R

The goal of this vignette is to present the protocol to *upload* data into the
`mangal` database. By default, the `rmangal` package works with the database
from the *Université du Québec à Rimouski*, in which data are publicly
available under the *Creative Commons 0 waiver*.

If you do not have an identifier, the first thing is to create one. This can
be done on line at `http://mangal.uqar.ca/dashboard/login/`. Once registered,
you will be given an *API key*. The recommended way to go from here is to
put this information in the `.Rprofile` file. The login function will either
take values passed as its `usr` and `key` arguments, or look for pre-set
`mangal_usr` and `mangal_key` values in the `.Rprofile`. For the rest of the
vignette, we will assume that you have stored your username and key in the
`.Rprofile` (and re-started your `R` session).

The next connection will automatically log you in:

```{r connect, eval=FALSE}
api <- mangalapi()
```

If you have *private* data, they will now appear in the output of `list`
and `get` commands.

# Upload walkthrough

As explained in the vignette on the basics, querying information is best done
from the top (`dataset` objects) to the bottom. The upload process follows the
*inverse* progression: first create taxa, then interactions, then networks,
then wrap them up in a dataset. It is important that you familiarize yourself
with the complete data specification *before* uploading your data.

## Creating taxa and related objects

## Creating datasets and related objects

## Releasing the dataset

The data you upload are *by default* public, and released under a *CC-0*
waiver. You can decide to make some of them private, simply by setting the
field `private` to `FALSE` for objects that have such a field. As per the
guidelines given on the website, you are expected not to treat the database
as a private data repository, and we expect that all deposited data will be
released in a reasonable amount of time. On your dashboard on the website,
you can check the status of your datasets and networks.

Releasing a dataset is done by calling a single function:
