<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Filtering in rmangal}
-->

```{r load,warning=FALSE, eval=TRUE, echo=FALSE, message=FALSE}
library(rmangal)
```
# Filtering

The `rmangal` package allows to *filter* ressources according to a series of
pre-defined relationships. Filtering is done in the same functions used to get
a list of objects, simply by passing an array of filters as the last argument.
The filters are lists, that have three attributes: `field`, `relation`, and
`target`. The basis of filtering is: the `field` of a given type of resource
will be matched for `target`, by a `relation`.

The `relation` can be any of `contains` (target matches anywhere in the field),
`startswith` or `endswith` (matches at the beginning or end of field), `exact`
(the field value is *exactly* target), `in` (target is one of the values of
the field), `range` (target is between the two bounds), and `gte`, `lte`,
`lt`, and `gt` (for greater/less or equal than).

By defaults, all fields are searchable (with the exception of `id` and
`owner`).

# Examples

```{r connec}
api <- mangalapi()
```

## Looking for all species of a given genus

To look for all species of  the genus `Heteractis`, we write the following
filter:

```{r spFilter}
heteractis <- list(field='name', relation='startswith', target='Heteractis')
heteractis_filter <- makefilter(api, 'taxa', heteractis)
Heteractis_species <- listTaxa(api, 'taxa', heteractis_filter)
laply(Heteractis_species, function(x) x$name)
```

The `makefilter` function will run several checks to ensure that the filter
is usable, and return a list with the informations needed for filtering.

## Looking for networks within a geographical boundary

We want to find networks within a square defined by its latitude/longitude
coordinates. We will define two boundaries: one for the latitudes, and one
for the longitudes. This can be done using `range`, and two filters:

```{r filtLatLon}
f_lat <- makefilter(api, 'network', list(field='latitude', relation='range', target=c(1.756, 1.813)))
f_lon <- makefilter(api, 'network', list(field='longitude', relation='range', target=c(124.760, 124.808)))
Networks <- listNetwork(api, c(f_lat, f_lon))
laply(Networks, function(x) x$name)
```

Note that in that situation, the several filters are combined with `c`.
