<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Uploading data}
-->

```{r load}
library(rmangal)
```

# Contributing data through R

The goal of this vignette is to present the protocol to *upload* data into the `mangal` database. By default, the `rmangal` package works with the database from the *Université du Québec à Rimouski*, in which data are publicly available under the *Creative Commons 0 waiver*.

If you do not have an identifier, the first thing is to create one:

```{r signUp, eval=FALSE}
me <- signUp(mangalapi(), 'my_username', 'my_password')
# Let's add some informations
me$first_name <- 'My first name'
me$last_name <- 'My last name'
me$email <- 'my@email.edu'
# We then connect using the new identifier
api <- mangalapi(usr=me$username, pwd=me$password)
# And finally, first thing first, let's update our contact info
me <- patchUser(api, me)
```

You only need to sign-up once. You can start your future sessions with

```{r login, eval=FALSE}
api <- mangalapi(usr='my_username',pwd='my_password')
```

# The upload approach

As explained in the vignette on the basics, querying information is best done from the top (`dataset` objects) to the bottom. The upload process follows the *inverse* progression: first create taxa, then interactions, then networks, then wrap them up in a dataset.

## Creating taxa objects

